---
title: "Arable Land vs Total Land Area by Country"
author: "Joel Schargorodsky"
date: "2024-05-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

# Arable Land vs Total Land Area by Country

Here, I will compile a list of all countries in the world, and will analyze the relationship between amount of arable land and total land area for each of these.

Name: Joel Schargorodsky

Email: joel.schargorodsky@gmail.com

# Part 1: A "jackknife" procedure to find the most outlying point in a linear relationship between two variables

First load the "XML" package to give access to `readHTMLTable()` and the "curl" package for access to `curl()`.
```{r}
if (!require("XML")) {
  install.packages("XML")
  stopifnot(require("XML"))
}
if (!require("curl")) {
  install.packages("curl")
  stopifnot(require("curl"))
}
```


Here, I will analyze the relationship between the land area and median home prices of each US state and Washington District of Columbia. I obtained the data from this webpage: https://en.wikipedia.org/wiki/List_of_U.S._states_by_median_home_price.

```{r}
link = "https://en.wikipedia.org/wiki/List_of_U.S._states_by_median_home_price"

tables = readHTMLTable(readLines(curl(link), warn=FALSE), skip.rows=c(1,3,26))


for (i in 1:3) {
  tables[[1]][,i] = noquote(sub(pattern=",", replacement="", x=tables[[1]][,i]))
}

state.index = tables[[1]][,2]
price.index = tables[[1]][,3]

indices = order(state.index)
price = price.index[indices]

for (i in 1:50) {
  priceString = toString(price[i])
  priceString = sub("$", "", priceString, fixed = TRUE)
  price[i] = as.numeric(priceString)
  print(price[i])
}

```

Use R to get farm areas of states from "population.csv".
```{r}
population.data = scan(file="populationTwo.csv", what="")

population.data = noquote(sub(pattern=",", replacement="", x = population.data))

state = population.data[seq(from=3, to=101, by=2)]
population = population.data[seq(from=4, to=102, by=2)]
population = as.numeric(population)
price = as.numeric(price)
```

Create a data frame called "affordability" whose columns are "state", "population", and "price", which contain state names, populations, and median house prices, respectively.
```{r}
 affordability <- data.frame(state,population,price)

```

Make a scatterplot of y = price vs. x = population.
```{r}
plot(x = population, y = price, xlab="population", ylab="price", xlim = c(0,40000000), ylim = c(0,1000000))
m = lm(formula = price~population, data=affordability)
abline(m, col = "red")
plot(m$residuals)
```
Note that in the plot above, one can see that 19 states have residuals with absolute value exceeding 100,000 and that four states have residuals with absolute value exceeding 200,000.

The states with greatest reisduals (say, those greater than 200000) are Hawaii, Montana, Massachusetts, and Washington. Hawaii has the greatest residual by far.

I will make a scatterplot again, this time with Hawaii removed.
```{r}
stateModuloHawaii = state[c(1:10, 12:50)]
populationModuloHawaii = population[c(1:10, 12:50)]
priceModuloHawaii = price[c(1:10, 12:50)]
plot(x = populationModuloHawaii, y = priceModuloHawaii, xlab="population", ylab="price", xlim = c(0,40000000), ylim = c(0,1000000))

affordabilityModuloHawaii <- data.frame(stateModuloHawaii,populationModuloHawaii,priceModuloHawaii)


mModuloHawaii = lm(formula = priceModuloHawaii~populationModuloHawaii, data=affordabilityModuloHawaii)
abline(mModuloHawaii, col = "red")
plot(mModuloHawaii$residuals)
```

Next we use a loop to create n=50 models. In step i, make a model of the data with observation i removed. Then predict the value of y[i] from that model, and find the residual (difference) between (the removed) y[i] and the prediction. Save these residuals in a vector `r.jack`. (A "jackknife" procedure works by removing one observation (or several) from a data set, and then making a prediction from that smaller data set, and repeating this for each observation.)
```{r}
state.aux = state
price.aux = price
population.aux = population
population.supp = population

r.jack = c()

for (i in 1:50) {
  ind = 1:50
  ind = ind[ind!= i]
  state.aux = state.aux[ind]
  price.aux = price.aux[ind]
  population.aux = population.aux[ind]
  m.i = lm(formula = population.aux ~ price.aux, data = affordability)
  population.supp[i] = m.i$coefficients[2]*price[i] + m.i$coefficients[1]
  r.jack = append(r.jack, population[i] - population.supp[i])
}
```

Plot these "jackknife" residuals.
```{r}
plot(r.jack)

```
It therefore turns out that only four states have "jackknife" residuals exceeding 100,000 and only two states with "jackknife" residuals exceeding 200,000. These are the only outlying states, unlike the more numerous set we encountered in our earlier residual plot.
